---
title: "Simple Geocoding Example"
author: "Mitchell J. Lovett"
date: "11/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

We will use the openstreetmaps API for this example. This is a public API. Note you can also use the Google maps API, but that requires more setup and has a licensing aspect to it.

```{r}
#install.packages("tmaptools")
#install.packages("here")
require(tmaptools)
require(here)
```

The task we will focus on is geocoding from a name and reverse coding from a lon-lat location. 
To get the location information for a point of interest, we need to pass the terms to search for: In this case we will pass TaiChi Bubble Tea, College Town, Rochester
```{r}
place1 = "Taichi Bubble Tea, College Town, Rochester"
location1 = geocode_OSM(place1,details = TRUE, as.data.frame = TRUE)
place2 = "Revolution Karaoke"
location2 = geocode_OSM(place2,details = TRUE, as.data.frame = TRUE)
location1
location2
```

This gives us the location of the retail restaurants!  

Now we reverse code this using just the lon and lat values. 

```{r}
placeReverse = rev_geocode_OSM(x = location1["lon"], y = location1["lat"],as.data.frame = TRUE)[, c("name")]
placeReverse

placeNearBy = rev_geocode_OSM(x = location1["lon"]-.0025, y = location1["lat"],as.data.frame = TRUE)[, c("name")]
placeNearBy
```
We can recover the Taichi place from its lat and lon, and we can "move" around by adjusting the lat and lon and seeing what we run into. . .

We can also evaluate distances between locations. We shouldn't use simple Euclidean distance and instead use so-called Haversine distance that deals with the spherical nature of the earth. 

```{r}
#Haversine distance - measures distance between points
# on "big circle" due to Earth's spherical shape
# see https://en.wikipedia.org/wiki/Haversine_formula
my_dist <- function(long1, lat1, long2, lat2) {
  rad <- pi/180
  a1 <- lat1*rad
  a2 <- long1*rad
  b1 <- lat2*rad
  b2 <- long2*rad
  dlon <- b2 - a2
  dlat <- b1 - a1
  a <- (sin(dlat/2))^2 + cos(a1)*cos(b1)*(sin(dlon/2))^2
  if(sqrt(a)>0 & sqrt(1-a)>0){
    c<- 2*atan(sqrt(a)/sqrt(1 - a))
  } else {
    c <- 2*atan2(sqrt(a), sqrt(1 - a))
  }
  R <- 6371/1.6 # radius in miles around Washington DC 
  d <- R*c
  return(d)
}

my_dist(location1["lon"],location1["lat"],location2["lon"],location2["lat"])

```

Get the 20m state boundary files from here:
https://www.census.gov/geographies/mapping-files/2013/geo/carto-boundary-file.html

```{r}
library(tidyverse)
library(sf)

usa <- "/Users/mitchlovett/Dropbox/Analytics Design/4. Geolocation Data and Segmentation/cb_2013_us_state_20m/cb_2013_us_state_20m.shp" %>%
  st_read()

(usa_48 <- usa %>%
  filter(!(NAME %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico"))))

load("/Users/mitchlovett/Box Sync/Mobile Case Data - Original/Simulated Data/MobileVisits.Rdata")

ggplot(data = usa_48) +
  geom_sf() + 
  geom_point(data = venues, aes(x = longitude, y = latitude,color = factor(chain)), alpha = .2) +
  coord_sf(xlim = c(-130, -60),
           ylim = c(20, 50))

```

