---
title: "yogurt"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#clean all data
```{r}
rm(list = ls())
```


## set up environment
```{r}
library(ggplot2)
library(gtools)
library(wordcloud)
library(tm)
library(reshape2)
library(foreign)

filnm <- "Wegmans Survey"
```

## read the survey
```{r}
survey <- read.spss(paste(filnm,".sav",sep=""),to.data.frame=TRUE,use.value.labels=FALSE)   #this one is numerical value
survey_factor <- read.spss(paste(filnm,".sav",sep=""), to.data.frame=TRUE,use.value.labels=TRUE,trim_values=TRUE)
#peek data
names(survey)

#summary
summary(survey_factor)
```

## extract performance variables
```{r}
perVar <- survey[,162:183]
names(perVar)
summary(perVar)
```


## ------- Face Validity Check ------- ##

```{r}
#check the marginal distribution graph
par(mfrow=c(3,3)); #specifies the arrangements of the plots
for (c in 1:length(survey_factor)) {
  x = survey_factor[,c]
  if (is.numeric(x)) {
    hist(x, main = names(survey_factor)[c]) #main specify the title
  } else {
    # plot(x, main = names(survey_factor)[c])
  }
}
```

## NA Check
```{r}
corCheck <- cor(survey_data, use = "pairwise.complete.obs")
sum(is.na(corCheck)/2)
summary(corCheck)
```


## check for perfect correlation
```{r}
#reset diagno to zero
diag(corCheck) = 0
sum(corCheck == 1, na.rm = TRUE) / 2
summary(corCheck == 1)
```

## free response check
```{r}
#questions: 14 16 17 18 19 23 26 29
free_response <- survey[c(78,87,89,91,92,103, 116, 129)]

#convert them to text
#what prepare with GY
prepare <- free_response[1][[1]]
#why choose GY over traditional
whyGY <- free_response[2][[1]]
#first reason for certian brand
reason1 <- free_response[3][[1]]
#why Fage
whyFage <- free_response[6][[1]]
#why Chobani
whyChobani <- free_response[7][[1]]
#why Oiko
whyOiko <- free_response[8][[1]]

txt <- list(prepare,whyGY,whyFage,whyChobani,whyOiko,reason1)

#use word cloud
for (i in 1:length(txt)) {
  docs = Corpus(VectorSource(txt[i]))
  #clean the text
  docs = tm_map(docs,removeNumbers)
  docs = tm_map(docs,removePunctuation)
  docs = tm_map(docs,stripWhitespace)
  docs = tm_map(docs,content_transformer(tolower))
  docs = tm_map(docs,removeWords,stopwords("english"))
  
  dtm = TermDocumentMatrix(docs)
  matrix = as.matrix(dtm)
  words = sort(rowSums(matrix),decreasing=TRUE) 
  df = data.frame(word = names(words),freq=words)
  
  #create the worldcloud
  set.seed(1234) #for repoducibility
  wordcloud(words=df$word,freq = df$freq, min.freq = 1,max.words = 100, random.order = FALSE,rot.per = 0.35,colors = brewer.pal(8,"Dark2"))
}
```

## Factor Analysis


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
